TINET+TECS: Component-Based TCP/IP Protocol Stack for Embedded Systems

### 2017-02-11 ISORC2017/paper.tex
Fix evaluation, related work, and conclusion by enago

This section describes our experimental evaluation to demonstrate the effectiveness of the proposed framework.

\subsection{Evaluation Environment}

GR-PEACH is employed as the evaluation board.
We connected the board and the host PC with a LAN cable, and evaluated the data transmission and reception.
The detail specification of the board is shown in TABLE \ref{tab:EvaluationBoardEnvironment}.
We also employ TINET 1.5.4 and the compiler arm-none-eabi--gcc 5.2

@@

To demonstrate the low overhead of TINET+TECS, we evaluated the execution time and the memory consumption of TINET+TECS compared with TINET.

The comparison of execution time between TINET and TINET+TECS is shown in Fig. \ref{fig:EvaluationOfExecutionTime}.
The APIs used for the evaluation are {\it tcp\_snd\_dat} to send TCP data and {\it tcp\_rcv\_dat} to receive TCP data.
For {\it tcp\_snd\_dat}, we measured the executing time from the API calling by the application to the return of the processing result.
In TINET+TECS, the process is performed in the order of tApplication, tTCPCEP, tTCPOutputTaskBody, tIPv4Output, tEthernetOutput, tArp, tEthernetOutputTaskBody, and tIfMbed of Fig. \ref{fig:ComponentProtocolStack}.
For {\it tcp\_rcv\_dat}, we measured the execution time from the data receiving in the LAN driver to the data acquisition in the application.
In TINET+TECS, the process is performed in the order of tIfMbed, tEthernetInputTaskBody, tIPv4Input, tTCPInput, tTCPCEP, and tApplication of Fig. \ref{fig:ComponentProtocolStack}.
The execution time of TINET+TECS is as well as that of TINET; the overhead is about 3 usec.
{\it send/receive} specifiers enable to access the buffer address without data copies.
Therefore, the overhead of componentization does not effect the execution time.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/EvaluationOfExecutionTime.pdf}
    \caption{Execution time of TINET and TINET+TECS}
    \label{fig:EvaluationOfExecutionTime}
\end{figure}

The memory consumptions of TINET and TINET+TECS are compared in TABLE \ref{tab:EvaluationOfMemoryConsumption}.
The memory consumption of TINET+TECS increases about 2.5\% compared with TINET.
The processes and data to manage TECS components, such as initialization of cells, descriptors, function tables, and skeleton functions, cause the increase memory consumption.

@@

As shown in TABLE \ref{tab:EvaluationOfConfigurabilty}, the code lines for modification were measured to demonstrate the improved configurability.
We can change the composition of the protocol stack with a small workload.
Thus, the proposed framework improves the configurability.

@@

\subsection{Dynamic Connection}

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/ComparisonOfDynamicConnection.pdf}
    \caption{Component diagrams of two cases (without/with dynamic connection)}
    \label{fig:ComparisonOfDynamicConnection}
\end{figure}

The memory consumption without and with TECS dynamic connection was evaluated.
As shown in the left of Fig. \ref{fig:ComparisonOfDynamicConnection}, a CEP component should statically connect to all REP components in the case dynamic connection is not used.
As the number of REPs increases, call ports of CEP are required for that.
Thus, it consumes a lot of memory. 
Dynamic connection reduces the memory consumption because only one call port of CEP for REP, which is drawn with red lines in the right of Fig. \ref{fig:ComparisonOfDynamicConnection}, is required per CEP.
Even when the number of REPs increases, a call port of the selector, not the CEP, is joined.

The memory consumption of two cases, without/with dynamic connection, is shown in Fig. \ref{fig:EvaluationOfDynamicConnection}.
The case of dynamic connection consumes the RAM memory more than the other case because the call ports with {\it [dynamic]} is not optimized and allocated in RAM areas as mentioned in Section \ref{sec:DynamicConnection}.
The total memory consumption is fewer in the proposed framework.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/EvaluationOfDynamicConnection.pdf}
    \caption{Memory consumption of two cases (with/without dynamic connection)}
    \label{fig:EvaluationOfDynamicConnection}
\end{figure}

@@

The comparison of CDL code lines of without/with dynamic connection is shown in TABLE \ref{tab:EvaluationOfConfigurabilityByDynamicConnection}, to demonstrate improved configurability by dynamic connection.
As the number of CEPs and REPs increases, the amount of CDL code lines to be added increases.
In the left part of Fig. \ref{fig:ComparisonOfDynamicConnection}, each CEP connects all REPs as shown in the upper of Fig. \ref{src:ComparisonOfCDL}. 
In the right part of Fig. \ref{fig:ComparisonOfDynamicConnection}, a CEP dynamically connects an REP, and only the selector connects all REPs as shown in the lower of Fig. \ref{src:ComparisonOfCDL}. 
Since the difference spreads as the number of CEPs and REPs increases, it is effective for software that uses many ports.

\subsection{Overhead of Adapter}

The execution time of API with the adapter such as {\it tcp\_snd\_dat/eAPI\_sendData} to analyze the overhead of TECS adapter which supports existing applications.
As shown in Fig. \ref{fig:EvaluationOfAdapter}, the overhead of TECS adapter is very small because the adapter only passes the parameters to TECS components from C codes.
Therefore, TECS adapter does not affect the system.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/EvaluationOfAdapter.pdf}
    \caption{Execution time of two cases (without/with TECS adapter)}
    \label{fig:EvaluationOfAdapter}
\end{figure}

@@

{\bf uIP:}
uIP (microIP) is a very small TCP/IP stack intended for tiny 8-bit and 16-bit microcontrollers.
    uIP only requires about five KB of code size and several hundreds bytes of RAM.
uIP has been ported to various systems and has found its way into many commercial products.
After ver. 1.0 is released, later versions of uIP, including uIPv6, are integrated with Contiki OS \cite{par:Contiki},\cite{url:Contiki}, an operating system to connect tiny microcontroller to the Internet.

{\bf lwIP:}
lwIP (lightweightIP) is a small TCP/IP implementation for embedded systems.
The focus of lwIP is to reduce memory resource while still having a full scale TCP.
lwIP requires about 40 KB of ROM and tens of KB of RAM.
lwIP is larger than uIP, but provides better throughput.

@@

This paper proposed TINET+TECS, a component-based TCP/IP protocol stack for embedded systems.
TINET+TECS is componentized TINET, which is a compact TCP/IP protocol stack, using TECS.
TINET consists of many macros and complicated codes and the software productivity are low.
The proposed framework improves the configurability while suppressing the overhead of componentization.
The proposed framework also improves the scalability because the component-based framework simplifies to add/remove and change protocols such as TCP/UDP, IPv4/IPv6, and Ethernet/PPP.

In addition, this paper presents dynamic connection, a new method of TECS, to enable dynamic processing while reducing memory consumption.
To satisfy TINET specification that TINET support static generation of CEPs and REPs, TINET+TECS utilizes dynamic connection.
TECS adapter supports legacy codes, existing TCP/IP applications can run without modification in the proposed framework.

In the future, the proposed framework will cooperate with mruby on TECS \cite{par:mrubyonTECS} to easily manage IoT devices.
Note that mruby is a scripting language for embedded systems \cite{par:mruby}.
We will support the functionalities that TINET functions can be utilized from mruby programs as an extension of mruby-socket \cite{url:mruby-library}.

>>>>

This section describes the experimental evaluation used to demonstrate the effectiveness of the proposed framework.

\subsection{Evaluation environment}

GR-PEACH was employed as the evaluation board.
Detailed specifications of the board are shown in TABLE \ref{tab:EvaluationBoardEnvironment}.
We also employ TINET 1.5.4 and the compiler arm-none-eabi--gcc 5.2
To pretest the system, we connected the board to a host PC via a LAN cable and evaluated the data transmission and reception.

@@

To demonstrate the low overhead of TINET+TECS, we compared its execution time and memory consumption with that of TINET, producing the comparison results shown in Fig. \ref{fig:EvaluationOfExecutionTime}.

The {\it tcp\_snd\_dat} and {\it tcp\_rcv\_dat} APIs were used in the evaluation to, respectively, send and receive TCP data.
For {\it tcp\_snd\_dat}, we measured the executing time starting from the API call through the application until the return of the processing result.
In TINET+TECS, this process is performed in the order tApplication, tTCPCEP, tTCPOutputTaskBody, tIPv4Output, tEthernetOutput, tArp, tEthernetOutputTaskBody, and tIfMbed, as shown in Fig. \ref{fig:ComponentProtocolStack}.
For {\it tcp\_rcv\_dat}, we measured the execution time from the data receipt in the LAN driver until data acquisition in the application.
In TINET+TECS, the process is performed in the order tIfMbed, tEthernetInputTaskBody, tIPv4Input, tTCPInput, tTCPCEP, and tApplication, as shown in Fig. \ref{fig:ComponentProtocolStack}.
The execution time of TINET+TECS is close to that of TINET, with an overhead of about 3 ms.
As the use of the {\it send/receive} specifier enables accessing of the buffer address without data copying, the componentization overhead does not affect the execution time.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/EvaluationOfExecutionTime.pdf}
    \caption{Execution times of TINET and TINET+TECS}
    \label{fig:EvaluationOfExecutionTime}
\end{figure}

The memory consumptions of TINET and TINET+TECS are compared in TABLE \ref{tab:EvaluationOfMemoryConsumption}.
The memory consumption of TINET+TECS is about 1\% higher than that of TINET, as the data and processes such as initialization of cells, descriptors, function tables, and skeleton functions needed to manage TECS components increase memory consumption.

@@

As shown in TABLE \ref{tab:EvaluationOfConfigurabilty}, the code lines for modification were measured to demonstrate the improved configurability.
This demonstrated the ability to change the composition of the protocol stack with a small workload, confirming that the proposed framework improves the configurability.

@@

\subsection{Dynamic connection}

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/ComparisonOfDynamicConnection.pdf}
    \caption{Component diagrams for two cases (without/with dynamic connection)}
    \label{fig:ComparisonOfDynamicConnection}
\end{figure}

Memory consumption without and with TECS dynamic connection was then evaluated.
As shown in the left panel of Fig. \ref{fig:ComparisonOfDynamicConnection}, each CEP component should be statically connected to all REP components if dynamic connection is not used.
As the number of REPs increases, additional call ports of CEP are required, in turn increasing the consumption of memory. 
Dynamic connection reduces memory consumption because only one CEP-to-REP call port is required per CEP, as illustrated with red lines in the right panel of Fig. \ref{fig:ComparisonOfDynamicConnection}.
Even if the number of REPs increases, additional call ports can be joined through the selector, instead of the CEPs.

Memory consumption of without and with dynamic connection is shown in Fig. \ref{fig:EvaluationOfDynamicConnection}.
The dynamic connection case consumes the more RAM memory because, as mentioned in Section \ref{sec:DynamicConnection}, call ports with {\it [dynamic]} are not optimized and allocated in RAM areas.
However, the overall memory consumption is lower under the proposed framework.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/EvaluationOfDynamicConnection.pdf}
    \caption{Memory consumption in two cases (with/without dynamic connection)}
    \label{fig:EvaluationOfDynamicConnection}
\end{figure}

@@

The comparison of CDL code lines of without and with dynamic connection is shown in TABLE \ref{tab:EvaluationOfConfigurabilityByDynamicConnection}, to demonstrate improved configurability by dynamic connection.
As the number of CEPs and REPs increases, the amount of CDL code lines to be added increases.
In the left panel of Fig. \ref{fig:ComparisonOfDynamicConnection}, each CEP connects all REPs as shown in the upper of Fig. \ref{src:ComparisonOfCDL}. 
In the right panel of Fig. \ref{fig:ComparisonOfDynamicConnection}, a CEP dynamically connects an REP, and only the selector connects all REPs as shown in the lower of Fig. \ref{src:ComparisonOfCDL}. 
It is effective for software that uses many ports because the difference spreads as the number of CEPs and REPs increases.

\subsection{Adapter overhead}

The API execution time when using an adapter such as {\it tcp\_snd\_dat/eAPI\_sendData} was used to analyze the overhead of the TECS adapter supporting existing applications.
As shown in Fig. \ref{fig:EvaluationOfAdapter}, the overhead of the TECS adapter is very small because the adapter only passes parameters from C codes to TECS components; thus, the TECS adapter overhead does not affect the system.

\begin{figure}[t]
    \centering
    \includegraphics[width=8.0cm,clip]{figure/EvaluationOfAdapter.pdf}
    \caption{Execution times in two cases (without/with TECS adapter)}
    \label{fig:EvaluationOfAdapter}
\end{figure}

@@

{\bf uIP:}
uIP (microIP) is a very small TCP/IP stack intended for tiny 8- and 16-bit microcontrollers.
uIP requires only about 5 KB of code size and several hundred bytes of RAM.
uIP has been ported to various systems and has found its way into many commercial products.
Following the release of ver. 1.0, later versions of uIP, including uIPv6, have been integrated with Contiki OS \cite{par:Contiki},\cite{url:Contiki}, an operating system to connect tiny microcontrollers to the Internet.

{\bf lwIP:}
lwIP (lightweightIP) is a small TCP/IP implementation for embedded systems that is intended to reduce memory resource usage while still maintaining a full-scale TCP.
lwIP requires about 40 KB of ROM and tens of KB of RAM.
lwIP is larger than uIP, but provides better throughput.

@@

This paper proposed TINET+TECS, a component-based TCP/IP protocol stack for embedded systems.
TINET+TECS is a componentized version of TINET, a compact TCP/IP protocol stack that uses TECS.
Because TINET comprises many macros and complicated codes, its software productivity is low.
The proposed framework improves on TINET's configurability while suppressing the overhead of componentization.
Scalability is also improved because the component-based framework simplifies to add/remove and change protocols such as TCP/UDP, IPv4/IPv6, and Ethernet/PPP.

This paper also presented dynamic connection, a new TECS method, to enable dynamic processing while reducing memory consumption.
TINET+TECS utilizes dynamic connection to satisfy the TINET specification for supporting static generation of CEPs and REPs.
As the TECS adapter supports legacy codes, existing TCP/IP applications can run without modification in the proposed framework.

In future work, we will adapt the proposed framework to cooperate with mruby on TECS \cite{par:mrubyonTECS} to more easily manage IoT devices.
Note that mruby is a scripting language for embedded systems \cite{par:mruby}.
We will support functionalities in which TINET functions can be utilized from mruby programs as an extension of mruby-socket \cite{url:mruby-library}.
